An Event-based Architecture for Multi-population Optimization Algorithms

### 2020-01-10 draft.tex
Addressing comment about async
Some comment about asynchrony would be in order

\begin{abstract} 
Multi-population based methods have been used extensively in
recent years to improve the performance of nature-inspired optimization
algorithms, these methods are parallel and asynchronous by nature and researchers
leverage this to scale algorithms % scale algorithms or implementations? - JJ
with distributed, multi-threaded,
parallel and cloud-based implementations. However, the design  of cloud-native,  distributed, multi-population algorithms is not a trivial task. Designers need to adapt the algorithms, while keeping its nature, to
leverage the scalability, elasticity, fault-tolerance,
reproducibility, and cost-effectiveness of reactive cloud-based systems. % reactive is mentioned here for the first time. Possibly you should just refer to "cloud-native" and then specify that it implies asynchrony, reactivity, and so on - JJ
In this
work, we present the design and implementation of a reactive container-based
application for the execution of multi-population algorithms. Researchers can
use the application to execute multiple experiments, both locally or in a cloud
environment and deploy experiments by specifying the infrastructure as part of
an experiment definition so that results are easily reproducible. We
evaluated the performance of the platform and the benefits of a multi-population
approach by using an ensemble of Genetic Algorithms (GAs) and Particle Swarm
Optimization (PSO) in a benchmark for the optimization of continuous functions.
Experiments show that the framework allows the ensemble to outperform
single-algorithm versions. The framework we provide also has fewer parameters to
tune since subpopulation parameters are selected randomly. The architecture and
the implemented platform is an excellent alternative for locally or a
cloud-based deployment.
\end{abstract}

@@

Population-based algorithms are intrinsically parallel, and can be implemented in an asynchronous. % They needn't be intrinsically asynchronous, I think. Most are implemented synchronouslu - JJ
The fitness of each individual can be evaluated independently of
others, and similarly, each population could evolve in isolation. Since 
earlier works, researchers have been proposing some form of parallelization
\cite{muhlenbein1988evolution} to increase the scalability of these algorithms.
The island model was one of the first techniques proposed for parallelization,
which lead to an increased performance \cite{gorges1990explicit,grosso1985computer}. 
The concept was to divide a large population into communicated subpopulations. 
Since then, other population-based algorithms have adopted the concept, 
and researchers have found additional benefits
besides the execution speed; these include avoiding a premature convergence and
maintaining the diversity of the global population. Researchers use the term
multi-population based methods when generally referring to techniques using
subpopulations as part of their optimization strategy. % Some comment about asynchrony would be in order - JJ

>>>>

\begin{abstract} 
Multi-population based methods have been used extensively in
recent years to improve the performance of nature-inspired optimization
algorithms, these methods are parallel and asynchronous by nature and researchers
leverage this to scale implementations % scale algorithms or implementations? - JJ   
%I think they are both, but in this case, implementation is more appropriate - Mario  
with distributed, multi-threaded,
parallel and cloud-based implementations. However, the design  of cloud-native,  
distributed, multi-population algorithms is not a trivial task. Designers need to adapt the algorithms, while keeping its nature, to
leverage the scalability, elasticity, fault-tolerance,
reproducibility, and cost-effectiveness of cloud-native systems. 
% reactive is mentioned here for the first time. Possibly you should just refer to "cloud-native" and then specify that it implies asynchrony, reactivity, and so on - JJ
% OK - Mario
In this
work, we present the design and implementation of a reactive container-based
application for the execution of multi-population algorithms. Researchers can
use the application to execute multiple experiments, both locally or in a cloud
environment and deploy experiments by specifying the infrastructure as part of
an experiment definition so that results are easily reproducible. We
evaluated the performance of the platform and the benefits of a multi-population
approach by using an ensemble of Genetic Algorithms (GAs) and Particle Swarm
Optimization (PSO) in a benchmark for the optimization of continuous functions.
Experiments show that the framework allows the ensemble to outperform
single-algorithm versions. The framework we provide also has fewer parameters to
tune since subpopulation parameters are selected randomly. The architecture and
the implemented platform is an excellent alternative for locally or a
cloud-based deployment.
\end{abstract}

@@

Population-based algorithms are intrinsically parallel, and can be implemented to run asynchronously. % They needn't be intrinsically asynchronous, I think. Most are implemented synchronouslu - JJ  
%OK - M
The fitness of each individual can be evaluated independently of 
others, and similarly, each population could evolve in isolation. Since 
earlier works, researchers have been proposing some form of parallelization
\cite{muhlenbein1988evolution} to increase the scalability of these algorithms.
The island model was one of the first techniques proposed for parallelization,
which lead to an increased performance \cite{gorges1990explicit,grosso1985computer}. 
The concept was to divide a large population into communicated subpopulations. 
Since then, other population-based algorithms have adopted the concept, 
and researchers have found additional benefits
besides the execution speed; these include avoiding a premature convergence and
maintaining the diversity of the global population. Researchers use the term
multi-population based methods when generally referring to techniques using
subpopulations as part of their optimization strategy. % Some comment about asynchrony would be in order - JJ
% Comments below - Mario
Moreover, parallel implementations can run in two ways: synchronous and
asynchronous. When operations run in parallel but maintain synchronism with
other operations,  all operations need to finish to continue to the next step.
For instance, in a  parent/worker model, the parent needs to wait for all
workers to finish their operations before moving to the next iteration.  In
contrast, in an asynchronous execution,  operations are not synchronized with
each other. In the previous, the parent operation can continue to the next
iteration, even if a worker has not finished its work.  Several works have
proposed asynchronous algorithms, reporting benefits in execution time, and
scalability. In these works, sub-populations
are processed independently and communicate asynchronously through a central
repository or message queue [refs needed to our previous work].
